---
title: "Corpora Exploratory Analysis"
author: "Refik Türkeli"
date: "12/8/2017"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("explore.R")
max_line <- -1
```

Source code for this document is available here: http://github.com/refik/capstone

## Project

Using mobile devices keyboards for inputting text to communicate or interact 
with applications takes significant time of device usage in daily life. The 
increase in processing power of mobile devices and advancements in artificial 
intelligence enables an opportunity for increasing the efficiancy of this 
process.

This project is about predictive text entry. I will create a prototype web
application using Shiny framework that will receive a stream of words by the 
user that is in the typing process and the application will try to predict
the next word that is intended to be written and show it as a couple of options
to the user. If one of the predictions is accurate, the user will select it. 

Accurate predictions will mean less time wasted on typing for the users. 

## Method

To predict the users next input, I will build a model that is trained by
text written previously by other people. This training will engrain the patterns
of text sequences into the model and enable it to generate predictions on a 
fresh stream of word input. 

The text to train the model with has to be ideally diverse to make the model
more successful in unique situations. Large sets of text data used for 
statistical analysis purposes are called *corpora*. 

## Corpora

The corpora that I will use for this project has text from `r length(language)` 
different languages and are obtained from `r length(type)` different contexts. 
I will highlight the features of the dataset with some data tables and figures.


#### Basic Statistics

There appears to be a balance for the amount of words available from each 
document type within a language. It is also not suprising to see that twitter
has the least words per line because of the 140 character limit. As a language,
English dominates the corpora considering the volume of text.

```{r, echo=FALSE, message=FALSE}
basic_stat_all(max_line = max_line) %>%
  basic_stat_kable()
```

#### Word Distribution for Documents

In the plot below we can see that news have a gaussian distribution accross
different languages. Meaning there is an avarage number of words for news 
which occurs most likely and as word count moves away from that average (more/less) 
there is less of them. It can be said that there is no distinctive 
characteristic for the amount of words in a news document and that its mostly random.

Blogs are different than news. Shortests blog documents occur most frequently
and increasing word count means less frequent occurance. My understanding is 
that people tend to write shorter blog posts.

```{r, echo=FALSE, message=FALSE, fig.align='center'}
word_count_all(max_line = max_line) %>%
  word_count_ggplot()
```

#### Most Frequent Words

I also analyzed the most frequent words in the corpora. Only English documents
are inspected for this purpose. Words like "the", "and" and "of" (stop words)
are filtered out.

Since the language is common, there are common words in the top 10 like "time",
"day" and "people". Twitter documents top 10 words have higher relative 
frequency than the other two types of documents. 

News have more serious words like "percent", "million" and "school" while
blogs have more lifestyle words like "world", "book" and "home". Twitters top 
words are more casual like "lol", "happy" and "night".

```{r, echo=FALSE, message=FALSE, fig.align='center'}
frequent_word_en(max_line = max_line) %>% 
  frequent_word_ggplot()
```

